---
output: html_document
editor_options: 
  chunk_output_type: console
---


#Load libraries
```{r}
# set session WD to source file location

library(dplyr)
library("gplots")
library(RColorBrewer)
library("ggplot2")
library("reshape2")
library("data.table")
library('randomcoloR')
library('ComplexHeatmap')
library('viridisLite')
library('openxlsx')
library('stringr')
library('lsa')
library('ggrepel')
library('VennDiagram')

library(msigdbr)

library(SummarizedExperiment)
library(ggplot2)
library(PerformanceAnalytics)
library(ggrepel)

library(UpSetR)
library(ggpubr)
library(ggcorrplot)
library(ggvenn)
library(VennDiagram)

library(data.table)
library(tibble)
library(purrr)
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)

library(viridis)
library(RColorBrewer)
```

#Functions and general 
```{r}
####### 
#General Functions
##########
#make all NA and NaN equal 0
is.nan.data.frame <- function(x){do.call(cbind, lapply(x, is.nan))}
is.inf.data.frame <- function(x){do.call(cbind, lapply(x, is.infinite))}

MakeNumeric <- function(mydata) {
  #mydata <- PGq
  mydata[is.nan.data.frame(mydata) == TRUE] <- 0 
  mydata[is.inf.data.frame(mydata) == TRUE] <- 0 
  mydata[mydata == 'NaN'] <- 0
  mydata[mydata == ''] <- 0
  mydata[is.na(mydata) == TRUE] <- 0 
  rownames(mydata) <- rownames(mydata)
  mydata2<- apply(mydata, 2, function(x){as.numeric(x)})
  mydata2 <<- mydata2
}

scale_row <- function(row) {
  (row - min(row)) / (max(row) - min(row))
}

theme_Publication <- function(base_size=16, base_family="helvetica") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family = "")
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "top",
            legend.direction = "horizontal",
            #legend.key.size= unit(0.2, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
  
}

my_palette5 <- brewer.pal(12,"Set3")
fill_color = c("#E2D200", "#94BF64", "#46ACC8")
#DD8D29 = control
#E2D200 = actin
#46ACC8 = arc
```

```{r}
date <- 'XXXXXXX'
```


#Load data 
```{r}

library(dplyr)
# Read files
pepDF <- read.csv('APEX_ProteinGroupReport.csv', header = T, stringsAsFactors = F)

# sample key
samplekey <- read.csv('APEX_sample_key.csv', header = T, stringsAsFactors = F)

IDs <- pepDF[,c(1:3)]
LFQs <- dplyr::select(pepDF, contains('PG.Label.Free.Quant'))
colnames(LFQs) <- paste0(samplekey$sample_n, '_', samplekey$sample_name, '_', samplekey$sample, '_', samplekey$rep.for.sample)
LFQs[LFQs == 'NaN'] <- NA

```

# number of IDs per sample
```{r}
colnames(LFQs) <-  c('NC_1', 'NC_2', 'Actb_1', 'Arc_1', 'NC_3', 'Actb_2', 'Arc_2')
LFQs <- LFQs[,c('NC_1', 'NC_2', 'NC_3', 'Arc_1', 'Arc_2', 'Actb_1', 'Actb_2')]


ids.perSample <- apply(LFQs, 2, function(x) {
  sum(!is.na(x))
})


pdf(paste0(date, '_IDsPerSample.pdf'), height = 5, width = 5)
barplot(ids.perSample)
dev.off()

```

# impute values around the 3rd percentile (add a pseudocount)
```{r}
##############
# imputation 
###############
lfq <- LFQs
# Find the next minimum value greater than 1
sorted_vals <- sort(LFQs[!is.na(LFQs)])
next_min_val <- sorted_vals[which(sorted_vals > 1)[1]]

# Calculate the bottom 3rd percentile value
third_percentile_val <- quantile(LFQs, 0.03, na.rm = TRUE)

# Generate a dataframe of the same dimensions as LFQs with values normally distributed
# between next_min_val and third_percentile_val

imputed_values <- as.data.frame(matrix(data = NA, nrow = nrow(LFQs), ncol = ncol(LFQs)))
for (i in 1:ncol(LFQs)) {
imputed_values[,i] <- rnorm(n = nrow(LFQs), mean = (next_min_val + third_percentile_val) / 2, 
                        sd = (third_percentile_val - next_min_val) / 6)
}
# Ensure no generated values are less than or equal to 1
while(any(imputed_values <= 1)) {
  imputed_values[imputed_values <= 1] <- rnorm(sum(imputed_values <= 1), 
                                               mean = (next_min_val + third_percentile_val) / 2, 
                                               sd = (third_percentile_val - next_min_val) / 6)
}

# Convert imputed values into a dataframe
LFQs_imputed <- LFQs
LFQs_imputed[is.nan.data.frame(LFQs_imputed) == T] <- 0
LFQs_imputed <- LFQs_imputed + imputed_values

# Display summary of the final data
summary(LFQs_imputed)

lfq <- LFQs_imputed

#####
# merge data with IDs
#####

loglfq <- cbind(IDs, log(lfq, 2))
lfq <- cbind(IDs, lfq)
write.csv(loglfq, file = paste0(date, '_logLFQ_doubleOnly_imputed.csv'))


```

# read previously imputed values
```{r}
#LFQs_imputed_log <- read.csv('20250407_logLFQ_doubleOnly_imputed.csv', header = T, stringsAsFactors = T)
# 20250407 is the worng file! I accidently overwritten on it on 20250521

loglfq_1 <- read.csv('APEXdata_imputed.csv', header = T, stringsAsFactors = T)

df <- loglfq_1[,-1]
LFQs <- df %>%
  mutate(across(where(is.numeric), ~ 2^.x))


loglfq <- loglfq_1[,-1]
```


# Differential abundance
## Fit linear model all samples
```{r}
library(limma)

# Create the design matrix

samplekey$sample <- (factor(samplekey$sample, levels = c('NC', 'actin', 'arc')))
design <- model.matrix(~0 + sample, data = samplekey)
colnames(design) <- levels(samplekey$sample)

# Fit the linear model
data <- MakeNumeric(loglfq[,-(1:3)])
#data <- MakeNumeric(LFQs_imputed)
fit <- lmFit(data, design) 

# Compute empirical Bayes statistics
fit <- eBayes(fit)

# set up contrasts
contrasts <- makeContrasts(
     arc_vs_control = arc - NC,
     act_vs_control = actin - NC,
     arc_vs_act = arc - actin,
     levels = design
 )

# Apply the contrasts to the fit object
fit2 <- contrasts.fit(fit, contrasts)
# 
# # Re-compute the empirical Bayes statistics for this specific contrast
fit2 <- eBayes(fit2)

results.adjusted <- summary(decideTests(fit2,
                    adjust.method = 'BH',
                    p.value = 0.1,
                    lfc = 0))
write.csv(results.adjusted, file = paste0(date, "_summary_differential_abundance.csv"))
```

## arc vs Control
```{r}
arc_vs_control <- topTable(fit2, coef="arc_vs_control", number = Inf, sort.by = 'none')
arc_vs_control <- cbind(arc_vs_control, IDs)

# write results
write.csv(arc_vs_control, file = paste0(date, "_arc_vs_control.csv"),
          row.names = FALSE)

```

# arc_vs_control volcano
```{r}

# volcano 
library("EnhancedVolcano")

# 1. Create a label list: top 20 genes with highest absolute logFC (regardless of sign)
top20_labels <- arc_vs_control |>
  dplyr::filter(!is.na(PG.Genes)) |>
  dplyr::arrange(desc(logFC)) |>
  dplyr::slice_head(n = 20) |>
  dplyr::pull(PG.Genes)

pdf(paste0(date, "_arc_vs_control_volcano.pdf"),
    width = 5,
    height = 5)
EnhancedVolcano(arc_vs_control,
                lab = arc_vs_control$PG.Genes,
                title = "Arc vs. Control",
                subtitle = NULL,
                caption = "Thresholds: FC = 4 ; adjusted p-value = 0.05",
                x = 'logFC',
                y = 'adj.P.Val',
                pCutoffCol = "adj.P.Val",
                pointSize = 3,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = log(4, 2),
                pCutoff = 0.05,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                drawConnectors = TRUE,
                colConnectors = 'black',
                ylim = c(0, 2.5),
                selectLab = top20_labels, # Creat a list of hits and supply here 
                maxoverlapsConnectors = 30 
                )
dev.off()

```

## act vs Control
```{r}
act_vs_control <- topTable(fit2, coef="act_vs_control", number = Inf, sort.by = 'none')
act_vs_control <- cbind(act_vs_control, IDs)

# write results
write.csv(act_vs_control, file = paste0(date, "_act_vs_control.csv"), row.names = FALSE)

```

# act vs Ctrl volcano
```{r}

library("EnhancedVolcano")

# 1. Create a label list: top 20 genes with highest absolute logFC (regardless of sign)
top20_labels <- act_vs_control |>
  dplyr::filter(!is.na(PG.Genes)) |>
  dplyr::arrange(desc(-log(adj.P.Val, 10))) |>
  dplyr::slice_head(n = 20) |>
  dplyr::pull(PG.Genes)

pdf(paste0(date, "_act_vs_control_volcano.pdf"),
    width = 5,
    height = 5)
EnhancedVolcano(act_vs_control,
                lab = act_vs_control$PG.Genes,
                title = "ACTB vs. Control",
                subtitle = NULL,
                caption = "Thresholds: FC = 4 ; adjusted p-value = 0.05",
                x = 'logFC',
                y = 'adj.P.Val',
                pCutoffCol = "adj.P.Val",
                pointSize = 3,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = log(4, 2),
                pCutoff = 0.05,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                drawConnectors = TRUE,
                colConnectors = 'black',
                ylim = c(0, 2.5),
                # xlim = c(-1.5, 1.5),
                selectLab = top20_labels, # Creat a list of hits and supply here 
                maxoverlapsConnectors = 30 
                )
dev.off()

```

##### Merge the significant lists
##### Plot Venn
##### Extract intersect and unique lists
```{r}

sig_act_0p1 <- act_vs_control[act_vs_control$adj.P.Val < 0.1 & act_vs_control$logFC > log(4,2),]
sig_arc_0p1 <- arc_vs_control[arc_vs_control$adj.P.Val < 0.1 & arc_vs_control$logFC > log(4,2),]


venn_analysis <- function(x, output_file = paste0(date, "_venn_analysis_results_limma.csv")) {
  # Function to get all combinations of a vector
  get_combinations <- function(vec) {
    unlist(lapply(seq_along(vec), function(i) combn(vec, i, simplify = FALSE)), 
           recursive = FALSE)
  }
  
  # Get all possible combinations of set names
  set_names <- names(x)
  combinations <- get_combinations(set_names)
  
  # Function to get the overlap of specific sets
  get_overlap <- function(sets) {
    Reduce(intersect, x[sets])
  }
  
  # Get overlaps
  overlaps <- lapply(combinations, get_overlap)
  
  # Name the overlaps
  names(overlaps) <- sapply(combinations, paste, collapse = " & ")
  
  # Get unique elements in each set
  unique_elements <- lapply(set_names, function(set) setdiff(x[[set]], unlist(x[set_names != set])))
  names(unique_elements) <- paste("Unique to", set_names)
  
  # Combine all results
  all_results <- c(overlaps, unique_elements)
  
  # Remove empty sets
  all_results <- all_results[sapply(all_results, length) > 0]
  
  # Create a data frame
  result_df <- data.frame(
    Category = names(all_results),
    Elements = sapply(all_results, paste, collapse = ", "),
    Count = sapply(all_results, length)
  )
  
  # Add columns for each original set
  for (set_name in set_names) {
    result_df[[set_name]] <- sapply(all_results, function(elements) {
      sum(elements %in% x[[set_name]])
    })
  }
  
  # Write to CSV
  write.csv(result_df, file = output_file, row.names = FALSE)
  
  #cat("Results have been written to", output_file, "\n")
  
  return(result_df)
}


x <- list(
"Actin vs control" = sig_act_0p1$PG.UniprotIds,
"Arc vs control" = sig_arc_0p1$PG.UniprotIds)


library(ggvenn)
pdf(paste0(date, "Venn_actin_vs_control_and_arc_vs_control.pdf"),
    width = 5,
    height = 5)
ggvenn(
  x, 
  fill_color = c("#40F0AE", "#F74ED6"),
  stroke_size = 0.5, set_name_size = 4
  )
dev.off()

#### Euler

library(eulerr)
# Create Euler fit (proportional Venn)
fit <- euler(x)

# Plot and save to PDF
pdf(paste0(date, "_Euler_actin_vs_control_and_arc_vs_control.pdf"),
    width = 5, height = 5)

plot(fit,
     fills = c("#40F0AE", "#F74ED6"),
     edges = TRUE,
     quantities = TRUE,
     labels = list(font = 1, cex = 1.2))

dev.off()

venn_analysis(x, paste0(date, "_arc_vs_control_and_actin_vs_control.csv"))
vennLists <- venn_analysis(x, paste0(date, "_arc_vs_control_and_actin_vs_control.csv"))


```

# Based on the Venn data
# subset the original data loglfq to the union of act and arc
```{r}
# Assuming vennLists$Elements is a string of comma-separated Uniprot IDs
venn_elements_act <- strsplit(vennLists$Elements, ",")[[1]]
venn_elements_arc <- strsplit(vennLists$Elements, ",")[[2]]
venn_elements_union <- union(venn_elements_act, venn_elements_arc)
venn_elements_arcUnique <- strsplit(vennLists$Elements, ",")[[5]]
venn_elements_intersect <- strsplit(vennLists$Elements, ",")[[3]]
venn_elements_actinUnique <- strsplit(vennLists$Elements, ", ")[[4]]

# Trim any whitespace (if any) from the elements
venn_elements_union <- trimws(venn_elements_union)
venn_elements_arcUnique <- trimws(venn_elements_arcUnique)
venn_elements_intersect <- trimws(venn_elements_intersect)


# convert to gene naems

map_uniprot_to_genes <- function(uniprot_ids, IDs) {
  # Check that required columns exist
  if (!all(c("PG.UniprotIds", "PG.Genes") %in% names(IDs))) {
    stop("IDs must contain columns: PG.UniprotIds and PG.Genes")
  }

  # Create a lookup vector
  id_map <- setNames(as.character(IDs$PG.Genes), IDs$PG.UniprotIds)

  # Map UniProt IDs to gene names
  gene_names <- id_map[uniprot_ids]

  # Return as character vector (with NAs if not found)
  return(gene_names)
} # for chr vectors

venn_elements_act_genes <- map_uniprot_to_genes(venn_elements_act, IDs)
venn_elements_arc_genes <- map_uniprot_to_genes(venn_elements_arc, IDs)
venn_elements_union_genes <- map_uniprot_to_genes(venn_elements_union, IDs)
venn_elements_arcUnique_genes <- map_uniprot_to_genes(venn_elements_arcUnique, IDs)
venn_elements_intersect_genes <- map_uniprot_to_genes(venn_elements_intersect, IDs)
venn_elements_actinUnique_genes <- map_uniprot_to_genes(venn_elements_actinUnique, IDs)

# write csv
write.csv(venn_elements_union_genes, file = paste0(date, '_vennLists_union.csv'))
write.csv(loglfq$PG.Genes, file = paste0(date, '_vennLists_Universe.csv'))
write.csv(venn_elements_arc_genes, file = paste0(date, '_vennLists_arcAll.csv'))
write.csv(venn_elements_act_genes, file = paste0(date, '_vennLists_actAll.csv'))
write.csv(venn_elements_actinUnique_genes, file = paste0(date, '_vennLists_act_unique.csv'))
write.csv(venn_elements_arcUnique_genes, file = paste0(date, '_vennLists_arc_unique.csv'))
write.csv(venn_elements_intersect_genes, file = paste0(date, '_vennLists_actArcIntersection.csv'))

# Filter the dataframe
df <- loglfq
filtered_loglfq <- df %>%
  filter(PG.UniprotIds %in% venn_elements_union)
  

```


######################################
# Check GO enrichment of the venn lits

# union of arc and act
```{r}

# Filter the dataframe
df <- loglfq
filtered_loglfq <- df %>%
  filter(PG.UniprotIds %in% venn_elements_union)


library('clusterProfiler')
library(org.Mm.eg.db)

enrich_up <- enrichGO(gene = as.character(filtered_loglfq$PG.Genes), # list of enriched genes
            universe = as.character(loglfq$PG.Genes), # all genes in the analysis
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "SYMBOL", #SYMBOL
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, paste0(date, "_act_arc_sig_union_GOall.csv"))

pdf(paste0(date, "_act_arc_sig_union_GOall.pdf"),
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "union arc and actin significant versus all Measured", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()

```
 
# arc separately
```{r}
library('clusterProfiler')
library(org.Mm.eg.db)

enrich_up <- enrichGO(gene = as.character(sig_arc_0p1$PG.Genes), # list of enriched genes
            universe = as.character(loglfq$PG.Genes), # all genes in the analysis
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "SYMBOL", #SYMBOL
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, file = paste0(date, "_arc_sig_GOall.csv"))

pdf(paste0(date,"_arc_sig_GOall.pdf"),
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "arc significant versus all Measured", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()


```

# act separately
```{r}
library('clusterProfiler')
library(org.Mm.eg.db)

enrich_up <- enrichGO(gene = as.character(sig_act_0p1$PG.Genes), # list of enriched genes
            universe = as.character(loglfq$PG.Genes), # all genes in the analysis
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "SYMBOL", #SYMBOL
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, file = (paste0(date,"_act_sig_GOall.csv")))

pdf(paste0(date,"_act_sig_GOall.pdf"),
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "actin significant versus all Measured", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()


```

# arc unique 
```{r}
library('clusterProfiler')
library(org.Mm.eg.db)

# Filter the dataframe
df <- loglfq
arc_unique <- df %>%
  filter(PG.UniprotIds %in% venn_elements_arcUnique)


enrich_up <- enrichGO(gene = as.character(arc_unique$PG.Genes), # list of enriched genes
            universe = as.character(loglfq$PG.Genes), # all genes in the analysis
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "SYMBOL", #SYMBOL
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, "arc_unique_GOall.csv")

pdf("arc_unique_GOall.pdf",
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "arc unique versus all Measured", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()


```

# arc act common
```{r}
library('clusterProfiler')
library(org.Mm.eg.db)

# Filter the dataframe
df <- loglfq
arc_act_common <- df %>%
  filter(PG.UniprotIds %in% venn_elements_intersect)


enrich_up <- enrichGO(gene = as.character(arc_act_common$PG.Genes), # list of enriched genes
            universe = as.character(loglfq$PG.Genes), # all genes in the analysis
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "SYMBOL", #SYMBOL
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, file = paste0(date, "_arc_act_common_GOall.csv"))

pdf(paste0(date, "_arc_act_common_GOall.pdf"),
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "intersection arc and actin significant versus all Measured", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()

```


###########################
# run the direct comparison on the loglfq values based on the union of arc and actin
```{r}

library(limma)

# Create the design matrix

samplekey$sample <- (factor(samplekey$sample, levels = c('NC', 'actin', 'arc')))
design <- model.matrix(~0 + sample, data = samplekey)
colnames(design) <- levels(samplekey$sample)

# define the data frame

df <- loglfq
filtered_loglfq <- df %>%
  filter(PG.UniprotIds %in% venn_elements_union)

data_exp <- MakeNumeric(filtered_loglfq[,-(1:3)])
data_IDs <- filtered_loglfq[,(1:3)]
data_IDs$PG.ProteinNames <- gsub('_MOUSE', '', data_IDs$PG.ProteinNames)
rownames(data_exp) <- data_IDs$PG.ProteinNames


fit <- lmFit(data_exp, design) 
#loglfq is the intensity values, log2 transformed and have to be in the same order as "samplekey".

# Compute empirical Bayes statistics
fit <- eBayes(fit)

# set up contrasts
contrasts <- makeContrasts(
     arc_vs_act = arc - actin,
     levels = design
 )

# Apply the contrasts to the fit object
fit2 <- contrasts.fit(fit, contrasts)

# # Re-compute the empirical Bayes statistics for this specific contrast
fit2 <- eBayes(fit2)

results.adjusted <- summary(decideTests(fit2,
                    adjust.method = 'BH',
                    p.value = 0.1,
                    lfc = 0))

arc_vs_act <- topTable(fit2, coef="arc_vs_act", number = Inf, sort.by = 'none')
subsetIDs <- filtered_loglfq[,1:3]

arc_vs_act <- cbind(arc_vs_act,
                        subsetIDs)


# write results
write.csv(arc_vs_act, paste0(date, "_arc_vs_act_direct_basedOnUnion.csv"),
          row.names = FALSE)

```


# volcano direct comparison
```{r}
pdf(paste0(date, "_arc_vs_act_direct_volcano.pdf"),
    width = 5,
    height = 5)
EnhancedVolcano(arc_vs_act,
                lab = arc_vs_act$PG.Genes,
                title = "Arc vs. Actin Direct (limma)",
                subtitle = NULL,
                caption = "Thresholds: FC = 2 ; p-value cutoff 0.05",
                x = 'logFC',
                y = 'P.Value',
                pCutoffCol = "P.Value",
                pointSize = 3,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = log2(2),
                pCutoff = 0.05,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                drawConnectors = TRUE,
                colConnectors = 'black',
                ylim = c(0, 5),
                selectLab = NULL, 
                maxoverlapsConnectors = 40
                )
dev.off()
```

# heatmap arc unique, act uniaue and intersection in 3 different plots

```{r}
# prepare the dataframes
sig_arc_direct <- arc_vs_act[arc_vs_act$P.Value < 0.05 & arc_vs_act$logFC > log(2,2),]
sig_act_direct <- arc_vs_act[arc_vs_act$P.Value < 0.05 & arc_vs_act$logFC < -log(2,2),]

# Combine all Uniprot IDs from the two filtered datasets to exclude from the union in order to define the "common"
excluded_ids <- unique(c(sig_arc_direct$PG.UniprotIds, sig_act_direct$PG.UniprotIds))

# Filter arc_vs_act to exclude those IDs
common_act_arc_direct <- arc_vs_act[!arc_vs_act$PG.UniprotIds %in% excluded_ids, ]


#########
# assign rownmaes to df (data matrix)
df <- filtered_loglfq
df <- df[, c(1:3, 6, 9, 7, 10, 4, 5, 8)]
data_IDs <- df[,(1:3)]
data_IDs$PG.ProteinNames <- gsub('_MOUSE', '', data_IDs$PG.ProteinNames)
rownames(df) <- data_IDs$PG.ProteinNames
```

```{r}
###################################################################

sig_arc_df <- filter(df, rownames(df) %in% rownames(sig_arc_direct))
df2plot_arc <- MakeNumeric(sig_arc_df[,-(1:3)])
rownames(df2plot_arc) <- rownames(sig_arc_df)

sig_act_df <- filter(df, rownames(df) %in% rownames(sig_act_direct))
df2plot_act <- MakeNumeric(sig_act_df[,-(1:3)])
rownames(df2plot_act) <- rownames(sig_act_df)

common_df <- filter(df, rownames(df) %in% rownames(common_act_arc_direct))
df2plot_comm <- MakeNumeric(common_df[,-(1:3)])
rownames(df2plot_comm) <- rownames(common_df)

library("ComplexHeatmap")


# Define annotation categories for columns (top bar)
column_annotation_labels <- case_when(
  grepl("arc", colnames(df2plot_comm)) ~ "arc",
  grepl("actin", colnames(df2plot_comm)) ~ "actin",
  grepl("NC", colnames(df2plot_comm)) ~ "NC",
  TRUE ~ "other"
)

column_annotation_colors <- c(
  "arc" = "#F74ED6", 
  "actin" = "#40F0AE", 
  "NC" = "gray"
)

# Create a top annotation object
column_ha <- HeatmapAnnotation(
  Condition = column_annotation_labels,
  col = list(Condition = column_annotation_colors)
)

pdf(paste0(date, "heatmap_actEnriched_inDirectAnalysis.pdf"),
    width = 5,
    height = 5)
Heatmap(t(scale(t(df2plot_act), center = TRUE, scale = TRUE)),
        column_title = "ACTB/ARC > 2",
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_column_names = FALSE,
        show_row_names = TRUE,
        cluster_column_slices = TRUE,
        cluster_row_slices = FALSE)
        #row_names_gp = gpar(fontsize = 3))
dev.off()

pdf(paste0(date, "heatmap_arcEnriched_inDirectAnalysis.pdf"),
    width = 5,
    height = 5)
Heatmap(t(scale(t(df2plot_arc), center = TRUE, scale = TRUE)),
        column_title = "ARC/ACTB > 2",
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_column_names = FALSE,
        show_row_names = TRUE,
        cluster_column_slices = TRUE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 7))
dev.off()

pdf(paste0(date, "heatmap_common_inDirectAnalysis.pdf"),
    width = 5,
    height = 5)
Heatmap(t(scale(t(df2plot_comm), center = TRUE, scale = TRUE)),
        column_title = "Not Differential",
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_column_names = FALSE,
        show_row_names = TRUE,
        cluster_column_slices = TRUE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 5))
dev.off()

write.csv(sig_act_direct, file = paste0(date, '_sig_act_direct.csv'))
write.csv(sig_arc_direct, file = paste0(date, '_sig_arc_direct.csv'))

```

# check lists against RBP2GO
```{r}
rbps_db <- read.csv('Table_RBP2GO.csv', header = T, stringsAsFactors = T)

library(dplyr)
rbps_in10ormore <- filter(rbps_db, as.numeric(Listing_Count) > 9)

# Create a function to count and calculate %
get_rbp_overlap_stats <- function(gene_list, rbp_genes) {
  overlap <- intersect(gene_list, rbp_genes)
  non_overlap <- setdiff(gene_list, rbp_genes)
  
  count_overlap <- length(overlap)
  percent_overlap <- (count_overlap / length(gene_list)) * 100
  
  return(list(
    count = count_overlap,
    percent = percent_overlap,
    overlapping_genes = overlap,
    non_overlapping_genes = non_overlap
  ))
}

# Apply to each list
stats_actin <- get_rbp_overlap_stats(venn_elements_actinUnique_genes, rbps_db$Gene_Name)
stats_arc   <- get_rbp_overlap_stats(venn_elements_arcUnique_genes, rbps_db$Gene_Name)
stats_inter <- get_rbp_overlap_stats(venn_elements_intersect, rbps_db$Gene_Name)
stats_union <- get_rbp_overlap_stats(venn_elements_union, rbps_db$Gene_Name)

# Print results
# Print summaries
cat("Actin unique:\n  Count:", stats_actin$count, 
    "\n  Percent:", round(stats_actin$percent, 2), "%\n",
    "  Overlapping Genes:\n", paste(stats_actin$overlapping_genes, collapse = ", "), "\n\n")

cat("Arc unique:\n  Count:", stats_arc$count, 
    "\n  Percent:", round(stats_arc$percent, 2), "%\n",
    "  Overlapping Genes:\n", paste(stats_arc$overlapping_genes, collapse = ", "), "\n\n")

cat("Intersection:\n  Count:", stats_inter$count, 
    "\n  Percent:", round(stats_inter$percent, 2), "%\n",
    "  Overlapping Genes:\n", paste(stats_inter$overlapping_genes, collapse = ", "), "\n\n")

cat("Union:\n  Count:", stats_union$count, 
    "\n  Percent:", round(stats_union$percent, 2), "%\n",
    "  Overlapping Genes:\n", paste(stats_union$overlapping_genes, collapse = ", "), "\n\n")
```


# Supplementary
#####################################
# bar plot sum intensity per sample
```{r}
df2plot <- loglfq[,4:10]

  df2plot_linear <- as.data.frame(lapply(df2plot, function(col) {
  if (is.numeric(col)) 2^col else col
}))
  
  
df2plot_linear[df2plot == 'NaN'] <- NA
#df2plot <- select(df2plot, -contains('cross'))
colnames(df2plot_linear) <- c("NC_1", "NC_2", "ACTB_1", "ARC_1", "NC_3", "ACTB_2", "ARC_2")
df2plot_order <- df2plot_linear[,c("NC_1", "NC_2", "NC_3",  "ARC_1", "ARC_2", "ACTB_1", "ACTB_2")]

sum_int <- apply(df2plot_order, 2, function(x){sum(x, na.rm = T)})

pdf(paste0(date, '_sumInt_perSample_LFQ_imputed_Linear.pdf'))
barplot(sum_int)
dev.off()

```


# scatters (SUP)
```{r}
library(ggplot2)

# Rename columns for convenience
data2plot <- loglfq
colnames(data2plot) <- c('UniprotIds', 'ProteinNames', 'Genes', 'Ctrl_1', 'Ctrl_2', 'Actb_1',
                         'Arc_1', 'Ctrl_3', 'Actb_2', 'Arc_2')

# Define pairs to plot
plot_pairs <- list(
  c('Ctrl_1', 'Ctrl_2'),
  c('Ctrl_2', 'Ctrl_3'),
  c('Ctrl_1', 'Ctrl_3'),
  c('Actb_1', 'Actb_2'),
  c('Actb_1', 'Ctrl_1'),
  c('Actb_1', 'Ctrl_2'),
  c('Actb_1', 'Ctrl_3'),
  c('Actb_2', 'Ctrl_1'),
  c('Actb_2', 'Ctrl_2'),
  c('Actb_2', 'Ctrl_3'),
  c('Arc_1', 'Arc_2'),
  c('Arc_1', 'Ctrl_1'),
  c('Arc_1', 'Ctrl_2'),
  c('Arc_1', 'Ctrl_3'),
  c('Arc_2', 'Ctrl_1'),
  c('Arc_2', 'Ctrl_2'),
  c('Arc_2', 'Ctrl_3')
)

# Loop through each pair and generate plot
for (pair in plot_pairs) {
  col1 <- pair[1]
  col2 <- pair[2]
  
  pdf(paste0(date, '_', col2, '_vs_', col1, '_scatter.pdf'), height = 5, width = 5)
  print(
    ggplot(data2plot, aes(x = .data[[col2]], y = .data[[col1]])) +
      geom_point(size = 2, alpha = 0.2) +
      #scale_fill_viridis_c() +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
      labs(
        #title = paste(col1, "vs", col2),
        x = col2,
        y = col1
      ) +
      theme_minimal() +
      theme(
      #plot.title = element_text(size = 14, face = "bold"),
      axis.title = element_text(size = 24, face = 'bold'),
      axis.text = element_text(size = 18, face = 'bold')  # axis tick labels

    )
  )
  dev.off()
}
```

