---
title: "Apex-Seq analysis - batch 2, Arc"
output: html_document
date: "2023-09-27"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes
- Batch 2 is the second library prep with higher RNA input and 16 cycles of amplification
- Following analysis as presented here: Padrón, Alejandro et al. “Proximity RNA Labeling by APEX-Seq Reveals the Organization of Translation Initiation Complexes and Repressive RNA Granules.” Molecular cell vol. 75,4 (2019): 875-887.e5. doi:10.1016/j.molcel.2019.07.030
- This analysis only analyzes the arc experimental samples


# Preparation

## Load libraries
```{r}
library(biomaRt)
library(msigdbr)

library(SummarizedExperiment)
library(DESeq2)
library(PCAtools)
library(genefilter)
library(zFPKM)
library(clusterProfiler)
library(org.Mm.eg.db)

library(ggplot2)
library(PerformanceAnalytics)
library(ggrepel)
library(EnhancedVolcano)
library(ComplexHeatmap)
library(UpSetR)
library(ggpubr)
library(ggcorrplot)
library(ggvenn)
#library(VennDiagram)

library(data.table)
library(tibble)
library(purrr)
library(dplyr)
library(tidyr)
library(reshape2)
library(stringr)

library(viridis)
library(RColorBrewer)
```

## Functions
```{r}
theme_Publication <- function(base_size=16, base_family="helvetica") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family = "")
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "top",
            legend.direction = "horizontal",
            #legend.key.size= unit(0.2, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
  
}

scale_fill_Publication <- function(...){
  library(scales)
  discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
  
}

scale_colour_Publication <- function(...){
  library(scales)
  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
  
}

venn_analysis <- function(x, output_file = "venn_analysis_results.csv") {
  # Function to get all combinations of a vector
  get_combinations <- function(vec) {
    unlist(lapply(seq_along(vec), function(i) combn(vec, i, simplify = FALSE)), 
           recursive = FALSE)
  }
  
  # Get all possible combinations of set names
  set_names <- names(x)
  combinations <- get_combinations(set_names)
  
  # Function to get the overlap of specific sets
  get_overlap <- function(sets) {
    Reduce(intersect, x[sets])
  }
  
  # Get overlaps
  overlaps <- lapply(combinations, get_overlap)
  
  # Name the overlaps
  names(overlaps) <- sapply(combinations, paste, collapse = " & ")
  
  # Get unique elements in each set
  unique_elements <- lapply(set_names, function(set) setdiff(x[[set]], unlist(x[set_names != set])))
  names(unique_elements) <- paste("Unique to", set_names)
  
  # Combine all results
  all_results <- c(overlaps, unique_elements)
  
  # Remove empty sets
  all_results <- all_results[sapply(all_results, length) > 0]
  
  # Create a data frame
  result_df <- data.frame(
    Category = names(all_results),
    Elements = sapply(all_results, paste, collapse = ", "),
    Count = sapply(all_results, length)
  )
  
  # Add columns for each original set
  for (set_name in set_names) {
    result_df[[set_name]] <- sapply(all_results, function(elements) {
      sum(elements %in% x[[set_name]])
    })
  }
  
  # Write to CSV
  write.csv(result_df, file = output_file, row.names = FALSE)
  
  cat("Results have been written to", output_file, "\n")
  
  return(result_df)
}

#' Make a paired heatmap of scaled expression and gene-set overlaps
#'
#' @param dds DESeq2 object with counts, rowData(external_gene_name), and colData(group, condition, replicate).
#' @param res_df Data.frame with columns 'external_gene_name', 'padj', 'log2FoldChange' used to order/filter rows.
#' @param previous_hits Named list of character vectors; each element is a gene set.
#' @param groups Character vector of sample groups to include (must match colData(dds)$group).
#' @param padj_cutoff Numeric FDR threshold. Default 0.05.
#' @param lfc_direction One of c("up","down","both"). Default "up".
#' @param lfc_min Minimum |log2FC| for filtering when lfc_direction != "both". Default 0.
#' @param sample_label_fun Function taking colData(dds) and returning character labels for columns.
#'        Default: paste(df$condition, df$replicate).
#' @param left_right_ratio Numeric length-2 vector giving width ratio for left:right heatmaps. Default c(2,1).
#' @param rowname_fontsize Numeric row name font size for left heatmap. Default 7.
#' @param pdf_file Optional output path; if NULL, no file is written. Default NULL.
#' @param pdf_width,pdf_height Numeric PDF size in inches if pdf_file is set. Defaults 6, 3.5.
#'
#' @return Invisibly returns a list with components: ht (CombinedHeatmap), n_sig (row count), genes (character).
#' @examples
#' # ht_info <- make_overlap_heatmap(dds, actin_vs_no_infect_scramble, previous_hits,
#' #                                 groups = c("control","actin"),
#' #                                 pdf_file = "actin_vs_no_infect_scramble_sig_heatmap.pdf")
make_overlap_heatmap <- function(
  dds,
  res_df,
  previous_hits,
  groups = c("control","actin"),
  padj_cutoff = 0.05,
  lfc_direction = "up",
  lfc_min = 0,
  sample_label_fun = function(df) paste(df$condition, df$replicate),
  left_right_ratio = c(2,1),
  rowname_fontsize = 7,
  pdf_file = NULL,
  pdf_width = 6,
  pdf_height = 3.5
) {
  stopifnot(all(c("external_gene_name","padj","log2FoldChange") %in% colnames(res_df)))
  stopifnot(length(left_right_ratio) == 2)

  # ---- libraries
  requireNamespace("DESeq2")
  requireNamespace("ComplexHeatmap")
  requireNamespace("circlize")
  requireNamespace("grid")

  # ---- normalized counts for selected groups
  keep_cols <- colData(dds)$group %in% groups
  if (!any(keep_cols)) stop("No columns match 'groups'.")
  mat <- as.matrix(DESeq2::counts(dds, normalized = TRUE)[, keep_cols, drop = FALSE])
  colnames(mat) <- sample_label_fun(as.data.frame(S4Vectors::DataFrame(colData(dds)))[keep_cols, , drop = FALSE])
  rownames(mat) <- S4Vectors::mcols(dds)$external_gene_name

  # ---- order to match res_df
  ord_idx <- match(res_df$external_gene_name, rownames(mat))
  ord_idx <- ord_idx[!is.na(ord_idx)]
  if (length(ord_idx) == 0) stop("No gene names in res_df match rownames(mat).")
  mat_order <- mat[ord_idx, , drop = FALSE]

  # ---- significance/filtering
  sig_idx <- !is.na(res_df$padj) & res_df$padj < padj_cutoff
  if (lfc_direction == "up")  sig_idx <- sig_idx & (res_df$log2FoldChange >  lfc_min)
  if (lfc_direction == "down") sig_idx <- sig_idx & (res_df$log2FoldChange < -lfc_min)
  # "both" ignores sign threshold besides padj
  sig_genes <- res_df$external_gene_name[sig_idx]
  sig_genes <- sig_genes[!is.na(sig_genes)]
  mat_sig <- mat_order[rownames(mat_order) %in% sig_genes, , drop = FALSE]
  # guard
  if (nrow(mat_sig) == 0) stop("No significant genes after filtering.")

  # ---- overlap matrix vs gene sets
  if (is.null(names(previous_hits))) names(previous_hits) <- paste0("Set", seq_along(previous_hits))
  ov_logical <- sapply(previous_hits, function(gs) rownames(mat_sig) %in% gs)
  ov_df <- as.data.frame(ov_logical, check.names = FALSE)

  # ---- scale per gene (row-wise z-score)
  mat_sig_scaled <- t(scale(t(mat_sig)))
  rownames(mat_sig_scaled) <- rownames(mat_sig)  # keep names if scale drops them

  # ---- sanity
  rownames(ov_df) <- rownames(mat_sig_scaled)
  stopifnot(identical(rownames(mat_sig_scaled), rownames(ov_df)))

  # ---- palettes
  overlap_cols <- c("TRUE" = "#2CA02C", "FALSE" = "ivory2")
  set_cols <- stats::setNames(grDevices::hcl.colors(ncol(ov_df), "Set2"), colnames(ov_df))

  # ---- categorical overlap strings
  ov_chr <- ifelse(as.matrix(ov_df), "TRUE", "FALSE")
  storage.mode(ov_chr) <- "character"
  dimnames(ov_chr) <- list(rownames(mat_sig_scaled), colnames(ov_df))

  # ---- annotations
  gs_factor <- factor(colnames(ov_chr), levels = colnames(ov_chr))
  top_anno_right <- ComplexHeatmap::HeatmapAnnotation(
    GeneSet = gs_factor,
    col = list(GeneSet = set_cols),
    which = "column",
    annotation_name_side = "left",
    show_legend = FALSE
  )

  cond_all <- as.character(colData(dds)$group)[keep_cols]
  cond_fac <- factor(cond_all, levels = unique(groups), labels = c("Ctrl","Arc")[seq_along(unique(groups))])
  ha <- ComplexHeatmap::HeatmapAnnotation(
    Condition = cond_fac,
    col = list(Condition = stats::setNames(c("grey", "#FF41C3")[seq_along(unique(groups))],
                                           levels(cond_fac)))
  )

  # ---- single legends
  lg_overlap <- ComplexHeatmap::Legend(
    title = "Overlap",
    at = c("TRUE","FALSE"),
    labels = c("TRUE","FALSE"),
    legend_gp = grid::gpar(fill = c("#2CA02C","ivory2"))
  )
  lg_sets <- ComplexHeatmap::Legend(
    title = "Gene sets",
    at = names(set_cols),
    legend_gp = grid::gpar(fill = unname(set_cols))
  )

  # ---- widths
  left_w  <- grid::unit(left_right_ratio[1], "null")
  right_w <- grid::unit(left_right_ratio[2], "null")

  # ---- heatmaps
  ht_left <- ComplexHeatmap::Heatmap(
    mat_sig_scaled,
    name = "Z-score",
    top_annotation = ha,
    border = TRUE,
    cluster_rows = TRUE,
    cluster_columns = TRUE,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    show_row_names = TRUE,
    row_names_side = "left",
    row_names_gp = grid::gpar(fontsize = rowname_fontsize),
    show_column_names = FALSE,
    width = left_w
  )

  ht_right <- ComplexHeatmap::Heatmap(
    ov_chr,
    name = "Overlap",
    col = overlap_cols,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    show_row_names = FALSE,
    show_column_names = FALSE,
    rect_gp = grid::gpar(col = NA),
    width = right_w,
    top_annotation = top_anno_right,
    show_heatmap_legend = FALSE
  )

  combined <- ht_left + ht_right

  # ---- draw (optionally to PDF)
  if (!is.null(pdf_file)) {
    grDevices::pdf(pdf_file, width = pdf_width, height = pdf_height)
    on.exit(try(grDevices::dev.off(), silent = TRUE), add = TRUE)
  }
  ComplexHeatmap::draw(
    combined,
    heatmap_legend_side = "right",
    annotation_legend_side = "right",
    annotation_legend_list = list(lg_overlap, lg_sets)
  )

  invisible(list(ht = combined,
                 n_sig = nrow(mat_sig_scaled),
                 genes = rownames(mat_sig_scaled)))
}

```

## Load previous data
```{r}
glock.rna <- read.csv("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/previous_studies/Glock et al 2021/rna-seq.csv")

caligas <- read.csv("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/previous_studies/Cajigas et al 2012/table S10.csv")

previous_hits <- list(
"Cajigas et al 2012 (Neuropil)" = caligas$Gene.Symbol,
"Glock et al 2021 (Neuropil)" = subset(glock.rna, rna_log2FC > 0 & rna_padj < 0.05)$gene,
"Glock et al 2021 (Somata)" = subset(glock.rna, rna_log2FC < 0 & rna_padj < 0.05)$gene)
```

## Load data
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis")

# metadata
metadata <- read.csv('sample_table.csv', header = TRUE, row.names = 1)

setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/results/archive/with-virus-and-stem-loop/archive")

tpm <- read.table('rsem.merged.gene_tpm.tsv', header = TRUE, row.names = 1)
tpm$transcript_id.s. <- NULL
colnames(tpm) <- sub("_(.*)$", "", colnames(tpm))
tpm <- tpm[,rownames(metadata)]


counts <- read.table('rsem.merged.gene_counts.tsv', header = TRUE, row.names = 1)
counts$transcript_id.s. <- NULL
colnames(counts) <- sub("_(.*)$", "", colnames(counts))
counts <- counts[,rownames(metadata)]

se <- SummarizedExperiment(assays=SimpleList(counts = as.matrix(round(counts)),
                                             tpm = as.matrix(tpm)),
                                             colData = metadata)

# remove version from gene name
rownames(se) <- lapply(rownames(se), sub, pattern = "\\.\\d+$", replacement = "")
```

## Subset to batch 2
```{r}
se <- se[, se$library.prep.batch == "2"]
```

## Subset to exclude libraries from batch 1 re-prepped in batch 2
```{r}
se <- se[, se$repeat.first.batch == "0"]
```

## Subset to exclude actin samples
```{r}
se <- se[, se$condition != "Actb"]
```

## Subset to exclude arc replicate 1
```{r}
se <- se[, !(se$condition == "arc" & se$replicate == 1)]
```

## Save
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed")

saveRDS(se, "se.RDS")
```

# Quality Control

## Raw genes expressed per sample
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

df <- data.frame(Sample = paste(se$sample.description, se$replicate), 
                 Frequency = colSums(assay(se, "tpm") > 1)) 

pdf("raw_counts_genes_expressed_tpm_1.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 1")
dev.off()

df <- data.frame(Sample = paste(se$sample.description, se$replicate), 
                 Frequency = colSums(assay(se, "tpm") > 5)) 

pdf("raw_counts_genes_expressed_tpm_5.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 5")
dev.off()

df <- data.frame(Sample = paste(se$sample.description, se$replicate), 
                 Frequency = colSums(assay(se, "tpm") > 10)) 

pdf("raw_counts_genes_expressed_tpm_10.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 10")
dev.off()
```

## Distribution of raw counts
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

df <- as.data.frame(log2(assay(se) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- gather(df, key = "Sample", value = "Value")

pdf("raw_count_distribution.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Value, fill = Sample)) +  # Use Value for density and Sample for color fill
    geom_density(alpha = 0.7) +  # Adjust alpha for transparency
    theme_Publication() +
    labs(x = "log2(Counts + 1)",  # Adjust the x-axis label for the density plot
         subtitle = "Raw Counts",
         y = "Density") +  # Density plots typically have "Density" as the y-axis label
    theme(legend.position = "right",
          legend.direction = "vertical")
dev.off()

df <- as.data.frame(log2(assay(se, "tpm") + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- gather(df, key = "Sample", value = "Value")

pdf("raw_tpm_distribution.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Value, fill = Sample)) +  # Use Value for density and Sample for color fill
    geom_density(alpha = 0.7) +  # Adjust alpha for transparency
    theme_Publication() +
    labs(x = "log2(TPM + 1)",  # Adjust the x-axis label for the density plot
         subtitle = "Raw TPM ",
         y = "Density") +  # Density plots typically have "Density" as the y-axis label
    theme(legend.position = "right",
          legend.direction = "vertical")
dev.off()
```

## Expression of stem loop sequences
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

stem_loop_seq <- c("24xMS2_gene", 
                  "24xpp7_gene", 
                  "actb_5-UTR_EXONS_3-UTR_24xms2_3-UTR_gene", 
                  "arc_5-UTR_EXONS_3-UTR_24xpp7v3_3-UTR_gene", 
                  "MS20_LK20_gene",
                  "MS2_LK51-1_gene",
                  "MS2_LK51-2_gene")

# counts
df <- as.data.frame(log2(assay(se) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_counts.pdf",
    width = 10,
    height = 10)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(counts)") +
  facet_wrap(~gene)
dev.off()

# tpm
df <- as.data.frame(log2(assay(se, "tpm") + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_tpm.pdf",
    width = 10,
    height = 10)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(tpm)") +
  facet_wrap(~gene)
dev.off()
```

### Filter arc and actb
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

stem_loop_seq <- c("actb_5-UTR_EXONS_3-UTR_24xms2_3-UTR_gene", 
                  "arc_5-UTR_EXONS_3-UTR_24xpp7v3_3-UTR_gene") 

# counts
df <- as.data.frame(log2(assay(se) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_counts-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(counts)") +
  facet_wrap(~gene)
dev.off()

# tpm
df <- as.data.frame(log2(assay(se, "tpm") + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_tpm-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(tpm)") +
  facet_wrap(~gene)
dev.off()
```

## Expression of stem loop sequences
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

stem_loop_seq <- c("24xMS2_gene", 
                  "24xpp7_gene", 
                  "actb_5-UTR_EXONS_3-UTR_24xms2_3-UTR_gene", 
                  "arc_5-UTR_EXONS_3-UTR_24xpp7v3_3-UTR_gene", 
                  "MS20_LK20_gene",
                  "MS2_LK51-1_gene",
                  "MS2_LK51-2_gene")

# counts
df <- as.data.frame(log2(assay(se) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_counts.pdf",
    width = 10,
    height = 10)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(counts)") +
  facet_wrap(~gene)
dev.off()

# tpm
df <- as.data.frame(log2(assay(se, "tpm") + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_tpm.pdf",
    width = 10,
    height = 10)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(tpm)") +
  facet_wrap(~gene)
dev.off()
```

## Expression of virus sequences
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

stem_loop_seq <- c("MCP_gene", 
                  "yGFP_gene", 
                  "PCP-APEX-P2A-yGFP_gene", 
                  "PCP_gene", 
                  "MCP-APEX-p2a-yGFP_gene")

# counts
df <- as.data.frame(log2(assay(se) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("virus_counts.pdf",
    width = 15,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(counts)") +
  facet_wrap(~gene, 
             nrow = 1)
dev.off()

# tpm
df <- as.data.frame(log2(assay(se, "tpm") + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("virus_tpm.pdf",
    width = 15,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(tpm)") +
  facet_wrap(~gene, 
             nrow = 1)
dev.off()
```

## Filter genes
- Keep genes which have at least 1 TPM in all arc samples
```{r}
se.filter <- se[rowSums(assay(se, "tpm")[,se$condition %in% c("arc")] > 1) == 2,]
```

## Filtered genes expressed per sample
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

df <- data.frame(Sample = paste(se.filter$sample.description, se.filter$replicate), 
                 Frequency = colSums(assay(se.filter, "tpm") > 1)) 

pdf("filtered_counts_genes_expressed_filterd_tpm_1.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 1")
dev.off()

df <- data.frame(Sample = paste(se.filter$sample.description, se.filter$replicate), 
                 Frequency = colSums(assay(se.filter, "tpm") > 5)) 

pdf("filtered_counts_genes_expressed_filterd_tpm_5.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 5")
dev.off()

df <- data.frame(Sample = paste(se.filter$sample.description, se.filter$replicate), 
                 Frequency = colSums(assay(se.filter, "tpm") > 10)) 

pdf("filtered_counts_genes_expressed_filterd_tpm_10.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Frequency)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(subtitle = "TPM > 10")
dev.off()
```

## Distribution of filtered counts
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

df <- as.data.frame(log2(assay(se.filter) + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- gather(df, key = "Sample", value = "Value")

pdf("filtered_count_distribution.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Value, fill = Sample)) +  # Use Value for density and Sample for color fill
    geom_density(alpha = 0.5) +  # Adjust alpha for transparency
    theme_Publication() +
    labs(x = "Log2(counts + 1)",  # Adjust the x-axis label for the density plot
         subtitle = "Post-filtering",
         y = "Density") +  # Density plots typically have "Density" as the y-axis label
    theme(legend.position = "right",
          legend.direction = "vertical")
dev.off()

df <- as.data.frame(log2(assay(se.filter, "tpm") + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- gather(df, key = "Sample", value = "Value")

pdf("filtered_tpm_distribution.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Value, fill = Sample)) +  # Use Value for density and Sample for color fill
    geom_density(alpha = 0.5) +  # Adjust alpha for transparency
    theme_Publication() +
    labs(x = "Log2(tpm + 1)",  # Adjust the x-axis label for the density plot
         subtitle = "Post-filtering",
         y = "Density") +  # Density plots typically have "Density" as the y-axis label
    theme(legend.position = "right",
          legend.direction = "vertical")
dev.off()
```

## Sample Correlation

### All samples
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

# tpm
df <- log2(assay(se.filter, "tpm") + 1)
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)

pdf("sample_correlation.pdf",
    width = 10,
    height = 10)
chart.Correlation(df, histogram=TRUE, pch=19)
dev.off()
```

### scrmbl
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

# tpm
df <- log2(assay(se.filter, "tpm") + 1)
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  as.data.frame() %>%
  select(contains("scramble"))

pdf("scramble_correlation.pdf",
    width = 5,
    height = 5)
chart.Correlation(df, histogram=TRUE, pch=19)
dev.off()
```

### ref transcriptome
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

# tpm
df <- log2(assay(se.filter, "tpm") + 1)
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  as.data.frame() %>%
  select(contains("transcriptome"))

pdf("ref-transcriptome_correlation.pdf",
    width = 5,
    height = 5)
chart.Correlation(df, histogram=TRUE, pch=19)
dev.off()
```

### arc
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

# tpm
df <- log2(assay(se.filter, "tpm") + 1)
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  as.data.frame() %>%
  select(contains("arc"))

pdf("acr_correlation.pdf",
    width = 5,
    height = 5)
chart.Correlation(df, histogram=TRUE, pch=19)
dev.off()
```

## Overlap analysis
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")

binary_expr <- as.data.frame(t(apply(assay(se.filter, "tpm"), 1, function(x) as.integer(x > 0))))
colnames(binary_expr) <- se.filter$sample.description

pdf("overlap_expressed_genes_all.pdf",
    width = 10,
    height = 7)
upset(binary_expr, 
      sets = colnames(binary_expr),
      text.scale = 1.25,
      #nintersects = 10,
      order.by = "freq",
      show.numbers = FALSE)
dev.off()

pdf("overlap_expressed_genes_top_10.pdf",
    width = 10,
    height = 7)
upset(binary_expr, 
      sets = colnames(binary_expr),
      text.scale = 1.25,
      nintersects = 10,
      order.by = "freq",
      show.numbers = FALSE)
dev.off()
```

## Map gene IDs
```{r}
# get database
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", 
                dataset = "mmusculus_gene_ensembl")

# mapping
names <- getBM(attributes = c('ensembl_gene_id', 
                              'external_gene_name', 
                              'entrezgene_id',
                              'gene_biotype',
                              'chromosome_name',
                              'start_position',
                              'end_position',
                              'percentage_gene_gc_content'),
                filters = 'ensembl_gene_id',
                values = rownames(se.filter),
                mart = mart)

# remove duplicates - 133
sum(duplicated(names$external_gene_name))
names <- names %>%
  distinct(external_gene_name, .keep_all = TRUE)

# match to genes in dataset
names.match <- names[match(rownames(se.filter), names$ensembl_gene_id),]

# add to se object row data
rowData(se.filter) <- names.match
all(rownames(rowData(se.filter)) == rowData(se.filter)$ensembl_gene_id)
```

## Save
```{r}
# counts
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/quality-control")
write.csv(merge(rowData(se.filter), assay(se.filter, "counts"), by = 'row.names'), "filtered.counts.csv")
write.csv(merge(rowData(se.filter), assay(se.filter, "tpm"), by = 'row.names'), "filtered.tpm.csv")

# object
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/")
saveRDS(se.filter, "se.filter.RDS")
se.filter <- readRDS("se.filter.RDS")
```

# Normalization
- Consider normalizing after taking out the virus and stem-loop sequences
- Consider normalizing to the amount of virus that is expressed
- Also consider normalizing without no infect controls as their pull down should be empty

## With virus and stem-loop sequences

### DeSeq2 normalization and modeling
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

# run deseq
dds <- DESeqDataSet(se.filter, design = ~condition)

dds <- DESeq(dds)
```

## Save and output normalized counts
```{r}
saveRDS(dds, "dds.RDS")
dds <- readRDS("dds.RDS")

dds.counts <- counts(dds, normalized = TRUE)
colnames(dds.counts) <- paste(dds$sample.description, dds$replicate, sep = " ")
dds.counts <- merge(dds.counts, rowData(dds), by = "row.names")

write.csv(dds.counts, "normalized.counts.csv")
```

## DESeq2 model estimates
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

# Mean and variance relationship
pdf("mean-variance.pdf",
    width = 5,
    height = 5)
norm.counts <- counts(dds, normalized=TRUE) 
mean.counts <- rowMeans(norm.counts)
variance.counts <- apply(norm.counts, MARGIN = 1, var)
mean.var.col <- densCols(x=log2(mean.counts), y=log2(variance.counts)) 
plot(x=log2(mean.counts), y=log2(variance.counts), pch=16, cex=0.5,
col=mean.var.col, main="Mean-variance relationship", xlab="Mean log2(normalized counts) per gene", ylab="Variance of log2(normalized counts)", panel.first = grid())
abline(a=1, b=1, col="red") # a linear line to compare against 
dev.off()

# dispersion
pdf("dispersion.pdf",
    width = 5,
    height = 5)
plotDispEsts(dds)
dev.off()
```

## Distribution of normalized counts
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

df <- as.data.frame(log2(counts(dds, normalized=TRUE) + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- gather(df, key = "Sample", value = "Value")

pdf("normalized_count_distribution_box.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = Sample, y = Value)) +
    geom_boxplot() +
    theme_Publication() +
    coord_flip() +
    labs(y = "Log2(counts + 1)",
         subtitle = "Normalized counts")
dev.off()

pdf("normalized_count_distribution_density.pdf",
    width = 6,
    height = 5)
ggplot(df, aes(x = Value, fill = Sample)) +
    geom_density(alpha = 0.5) +
    theme_Publication() +
    labs(x = "Log2(counts + 1)",
         subtitle = "Normalized counts") +
    theme(legend.position = "right",
          legend.direction = "vertical")
dev.off()
```

## Sample correlations
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

df <- as.data.frame(log2(counts(dds, normalized=TRUE) + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df.cor <- cor(df)
df.cor <- df.cor[order(rownames(df.cor)), order(colnames(df.cor))]

pdf("sample_correlations.pdf",
    width = 8,
    height = 8)
ggcorrplot(df.cor, 
           hc.order = FALSE, 
           type = "lower",
           lab = TRUE)
dev.off()
```

## Clustering
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

# Transform counts to stabalize mean-variance
dds_vst <- vst(dds, blind = FALSE)

# PCA
dds$replicate <- factor(dds$replicate)
p <- pca(assay(dds_vst), metadata = colData(dds), removeVar = 0.1)

pdf("screeplot.pdf",
    width = 5,
    height = 5)
screeplot(p)
dev.off()

pdf("eingencor.pdf",
    width = 5,
    height = 5)
eigencorplot(p,
             components = getComponents(p, seq_len(7)),
             main = "Principle component correlations",
             cexMain = 1.5,
             metavars = c("condition", "replicate"),
             col = viridis(100),
             colCorval = 'firebrick',
             fontCorval = 2,
             cexCorval = 0.5,
             rotLabX = 45,
             posColKey = 'top')
dev.off()

pdf("PCA.pdf",
    width = 6,
    height = 5)
biplot(p,
       colby = "condition",
       shape = "replicate",
       # hline = 0, 
       # vline = 0,
       gridlines.major = FALSE, 
       gridlines.minor = FALSE,
       max.overlaps = Inf,
       lab = dds$sample.description) +
       theme_Publication() +
       theme(legend.position = "right",
             legend.box="vertical",
             legend.direction = "vertical",
             legend.margin=margin())
dev.off()

pdf("PCA_no_labels.pdf",
    width = 6,
    height = 5)
biplot(p,
       colby = "condition",
       shape = "replicate",
       hline = 0, 
       vline = 0,
       gridlines.major = FALSE, 
       gridlines.minor = FALSE,
       lab = NULL) +
       theme_Publication() +
       theme(legend.position = "right",
             legend.box="vertical",
             legend.direction = "vertical",
             legend.margin=margin())
dev.off()
```

### Without background samples
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

# Transform counts to stabalize mean-variance
dds_vst <- vst(dds, blind = FALSE)

# remove background samples
dds_vst.sub <- dds_vst[,dds_vst$condition != "background"]

# PCA
dds_vst.sub$replicate <- factor(dds_vst.sub$replicate)
p <- pca(assay(dds_vst.sub), metadata = colData(dds_vst.sub), removeVar = 0.1)

pdf("screeplot_background_removed.pdf",
    width = 5,
    height = 5)
screeplot(p)
dev.off()

pdf("eingencor_background_removed.pdf",
    width = 5,
    height = 5)
eigencorplot(p,
             components = getComponents(p, seq_len(5)),
             main = "Principle component correlations",
             cexMain = 1.5,
             metavars = c("condition", "replicate"),
             col = viridis(100),
             colCorval = 'firebrick',
             fontCorval = 2,
             cexCorval = 0.5,
             rotLabX = 45,
             posColKey = 'top')
dev.off()

pdf("PCA_background_removed.pdf",
    width = 6,
    height = 5)
biplot(p,
       colby = "condition",
       shape = "replicate",
       # hline = 0, 
       # vline = 0,
       gridlines.major = FALSE, 
       gridlines.minor = FALSE,
       max.overlaps = Inf,
       lab = dds_vst.sub$sample.description) +
       theme_Publication() +
       theme(legend.position = "right",
             legend.box="vertical",
             legend.direction = "vertical",
             legend.margin=margin())
dev.off()

pdf("PCA_no_labels.pdf",
    width = 6,
    height = 5)
biplot(p,
       colby = "condition",
       shape = "replicate",
       hline = 0, 
       vline = 0,
       gridlines.major = FALSE, 
       gridlines.minor = FALSE,
       lab = NULL) +
       theme_Publication() +
       theme(legend.position = "right",
             legend.box="vertical",
             legend.direction = "vertical",
             legend.margin=margin())
dev.off()
```

## Plot expression of genes

### Virus sequences
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

stem_loop_seq <- c("yGFP_gene", 
                  "PCP-APEX-P2A-yGFP_gene", 
                  "MCP-APEX-p2a-yGFP_gene")

# counts
df <- as.data.frame(counts(dds, normalized=TRUE))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("virus_norm_counts.pdf",
    width = 10,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "normalized counts") +
  facet_wrap(~gene, 
             nrow = 1)
dev.off()

# counts log2
df <- as.data.frame(log2(counts(dds, normalized=TRUE) + 1))
colnames(df) <- paste(se$sample.description, se$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("virus_norm_counts_log2.pdf",
    width = 10,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(normalized counts)") +
  facet_wrap(~gene, 
             nrow = 1)
dev.off()
```

### Stem loop sequences
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

stem_loop_seq <- c("actb_5-UTR_EXONS_3-UTR_24xms2_3-UTR_gene", 
                  "arc_5-UTR_EXONS_3-UTR_24xpp7v3_3-UTR_gene") 

# counts
df <- as.data.frame(counts(dds, normalized=TRUE))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_norm_counts-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Normalized counts") +
  facet_wrap(~gene)
dev.off()


# counts log2
df <- as.data.frame(log2(counts(dds, normalized=TRUE) + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% stem_loop_seq) %>%
  mutate(gene = gsub("_gene", "", gene)) %>%
  mutate(gene = gsub("_", " ", gene))

df$gene <- str_wrap(df$gene, width = 15)

pdf("stem-loop_norm_counts_log2-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(normalized counts)") +
  facet_wrap(~gene)
dev.off()
```

### Endogenous actin and arc
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/normalization")

endogenous <- c("ENSMUSG00000029580", 
                "ENSMUSG00000022602")

gene_names <- c("ENSMUSG00000029580" = "Actb", 
                "ENSMUSG00000022602" = "Arc")

# counts
df <- as.data.frame(counts(dds, normalized=TRUE))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% endogenous) %>%
  mutate(gene_name = gene_names[gene])

df$gene <- str_wrap(df$gene, width = 15)

pdf("endogenous_norm_counts-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Normalized counts") +
  facet_wrap(~gene_name)
dev.off()


# counts log2
df <- as.data.frame(log2(counts(dds, normalized=TRUE) + 1))
colnames(df) <- paste(se.filter$sample.description, se.filter$replicate)
df <- df %>%
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = -gene, names_to = "sample", values_to = "value") %>%
  filter(gene %in% endogenous) %>%
  mutate(gene_name = gene_names[gene])

df$gene <- str_wrap(df$gene, width = 15)

pdf("endogenous_norm_counts_log2-arc-actb.pdf",
    width = 8,
    height = 5)
ggplot(df, aes(x = sample, y = value)) +
    geom_col() + 
    theme_Publication() +
    coord_flip() +
    labs(x = "",
         y = "Log2(normalized counts)") +
  facet_wrap(~gene_name)
dev.off()
```


# Differential Expression

## No infect vs background (ref transcriptome)
- Control for effect of pull down
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")

# get results
noinfect_vs_background <- results(dds, 
                  alpha = 0.05, 
                  contrast = c("condition", "no infect", "background"))

# get a summary of the amount of significantly DEG 
summary(noinfect_vs_background, 
        alpha = 0.05)

# get shrunken fold change
# noinfect_vs_background <- lfcShrink(dds = dds, 
#                               res = noinfect_vs_background, 
#                               contrast = c("condition", "no infect", "background"),
#                               type = "ashr")

cols <- c("ensembl_gene_id",
          "external_gene_name",
          "entrezgene_id",
          "gene_biotype",
          "chromosome_name",
          "start_position",
          "end_position",
          "percentage_gene_gc_content")

noinfect_vs_background <- merge(as.data.frame(rowData(dds)[,cols]),
                         as.data.frame(noinfect_vs_background),
                         by = "row.names")

counts <- counts(dds, normalized = TRUE)
colnames(counts) <- dds$sample.description
noinfect_vs_background <- merge(noinfect_vs_background, 
                         counts,
                         by.x = "Row.names",
                         by = "row.names")

noinfect_vs_background$ensembl_gene_id <- NULL
colnames(noinfect_vs_background)[1] <- "ensembl_gene_id"

# write results
write.csv(noinfect_vs_background, "noinfect_vs_background.csv",
          row.names = FALSE)

# p-value distribution histogram
pdf("noinfect_vs_background_pvalue_distribution.pdf",
    width = 5,
    height = 5)
hist(noinfect_vs_background$pvalue, 
     main = "P-value distribution", 
     xlab = "P-value", 
     ylab = "Frequency")
dev.off()


# ma plot
noinfect_vs_background <- noinfect_vs_background %>%
  as.data.frame() %>%
  mutate(sig = case_when(log2FoldChange > 0 & noinfect_vs_background$padj < 0.05 ~ "sig.up",
                         log2FoldChange < 0 & noinfect_vs_background$padj < 0.05 ~ "sig.down",
                         .default = "")
         )

pdf("noinfect_vs_background_ma_log2.pdf",
    width = 5,
    height = 5)
ggplot(noinfect_vs_background, aes(log2(baseMean + 1), log2FoldChange, color = sig)) +
  geom_point(alpha = 0.5, size = 4) +
  theme_Publication() +
  scale_colour_manual(values = c("grey", "blue", "red")) +
  geom_hline(yintercept= 0, 
             size=1,
             linetype = "dashed") +
  labs(title = "No infect vs Background",
       x = "Average Normalized Log2(counts) + 1",
       y = "Log2(fold change)")
dev.off()

# volcano plot
noinfect_vs_background$volcano.label <- ""
noinfect_vs_background$volcano.score <- abs(noinfect_vs_background$log2FoldChange) * -log10(noinfect_vs_background$pvalue)
noinfect_vs_background <- noinfect_vs_background[order(noinfect_vs_background$volcano.score, decreasing = TRUE),]
noinfect_vs_background$volcano.label[1:10] <- noinfect_vs_background$external_gene_name[1:10]

pdf("noinfect_vs_background_volcano.pdf",
    width = 5,
    height = 5)
EnhancedVolcano(noinfect_vs_background,
                lab = noinfect_vs_background$external_gene_name,
                title = "No infect vs Background",
                subtitle = NULL,
                caption = "FC > 2; padj < 0.05",
                x = 'log2FoldChange',
                y = 'pvalue',
                pointSize = 2,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = 2,
                pCutoff = 0.05,
                pCutoffCol = 'padj',
                selectLab = noinfect_vs_background$volcano.label,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = TRUE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                max.overlaps = Inf,
                gridlines.major = FALSE,
                gridlines.minor = FALSE)
dev.off()

# heatmap of significant changes
ha <- HeatmapAnnotation(Condition = subset(colData(dds), condition %in% c("background", "no infect"))$condition,
                        Replicate = as.factor(subset(colData(dds), condition %in% c("background", "no infect"))$replicate))

# get normalized counts
mat <- as.matrix(counts(dds, normalized = TRUE)[,colData(se.filter)$condition %in% c("background", "no infect")])

# order by results
mat.order <- mat[match(noinfect_vs_background$ensembl_gene_id, rownames(mat)),]

# get significant genes
mat.sig <- mat.order[noinfect_vs_background$padj < 0.05,]
mat.sig <- mat.sig[!is.na(row.names(mat.sig)),]

# scale and center normalized counts
mat.sig.scaled <- t(scale(t(mat.sig)))

pdf("noinfect_vs_background_sig_heatmap.pdf",
    width = 4,
    height = 5)
Heatmap(mat.sig.scaled, 
        name = "Z-score",
        top_annotation = ha,
        border = TRUE,
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_row_names = FALSE,
        show_column_names = FALSE,
        row_dend_width = unit(20, "mm"))
dev.off()
```

## Scramble vs background
- Control for effect of infection
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")

# get results
scramble_vs_background <- results(dds, 
                  alpha = 0.05, 
                  contrast = c("condition", "scramble", "background"))

# get a summary of the amount of significantly DEG 
summary(scramble_vs_background, 
        alpha = 0.05)

# get shrunken fold change
# scramble_vs_background <- lfcShrink(dds = dds, 
#                               res = scramble_vs_background, 
#                               contrast = c("condition", "background", "background"),
#                               type = "ashr")

# merge
cols <- c("ensembl_gene_id",
          "external_gene_name",
          "entrezgene_id",
          "gene_biotype",
          "chromosome_name",
          "start_position",
          "end_position",
          "percentage_gene_gc_content")

scramble_vs_background <- merge(as.data.frame(rowData(dds)[,cols]),
                         as.data.frame(scramble_vs_background),
                         by = "row.names")


counts <- counts(dds, normalized = TRUE)
colnames(counts) <- dds$sample.description
scramble_vs_background <- merge(scramble_vs_background, 
                         counts,
                         by.x = "Row.names",
                         by = "row.names")

scramble_vs_background$ensembl_gene_id <- NULL
colnames(scramble_vs_background)[1] <- "ensembl_gene_id"

# write results
write.csv(scramble_vs_background, "scramble_vs_background.csv",
          row.names = FALSE)

# p-value distribution histogram
pdf("scramble_vs_background_pvalue_distribution.pdf",
    width = 5,
    height = 5)
hist(scramble_vs_background$pvalue, 
     main = "P-value distribution", 
     xlab = "P-value", 
     ylab = "Frequency")
dev.off()


# ma plot
scramble_vs_background <- scramble_vs_background %>%
  as.data.frame() %>%
  mutate(sig = case_when(log2FoldChange > 0 & scramble_vs_background$padj < 0.05 ~ "sig.up",
                         log2FoldChange < 0 & scramble_vs_background$padj < 0.05 ~ "sig.down",
                         .default = "")
         )

pdf("scramble_vs_background_ma_log2.pdf",
    width = 5,
    height = 5)
ggplot(scramble_vs_background, aes(log2(baseMean + 1), log2FoldChange, color = sig)) +
  geom_point(alpha = 0.5, size = 4) +
  theme_Publication() +
  scale_colour_manual(values = c("grey", "blue", "red")) +
  geom_hline(yintercept= 0, 
             size=1,
             linetype = "dashed") +
  labs(title = "scramble vs Background",
       x = "Average Normalized Log2(counts) + 1",
       y = "Log2(fold change)")
dev.off()

# volcano plot
scramble_vs_background$volcano.label <- ""
scramble_vs_background$volcano.score <- abs(scramble_vs_background$log2FoldChange) * -log10(scramble_vs_background$pvalue)
scramble_vs_background <- scramble_vs_background[order(scramble_vs_background$volcano.score, decreasing = TRUE),]
scramble_vs_background$volcano.label[1:10] <- scramble_vs_background$external_gene_name[1:10]

pdf("scramble_vs_background_volcano.pdf",
    width = 5,
    height = 5)
EnhancedVolcano(scramble_vs_background,
                lab = scramble_vs_background$external_gene_name,
                title = "Scramble vs Background",
                subtitle = NULL,
                caption = "FC > 2; padj < 0.05",
                x = 'log2FoldChange',
                y = 'pvalue',
                pointSize = 2,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = 2,
                pCutoff = 0.05,
                pCutoffCol = 'padj',
                selectLab = scramble_vs_background$volcano.label,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = TRUE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                max.overlaps = Inf,
                gridlines.major = FALSE,
                gridlines.minor = FALSE)
dev.off()

# heatmap of significant changes
ha <- HeatmapAnnotation(Condition = subset(colData(dds), condition %in% c("background", "scramble"))$condition,
                        Replicate = subset(colData(dds), condition %in% c("background", "scramble"))$replicate)

# get normalized counts
mat <- as.matrix(counts(dds, normalized = TRUE)[,colData(se.filter)$condition %in% c("background", "scramble")])

# order by results
mat.order <- mat[match(scramble_vs_background$ensembl_gene_id, rownames(mat)),]

# get significant genes
mat.sig <- mat.order[scramble_vs_background$padj < 0.05,]
mat.sig <- mat.sig[!is.na(row.names(mat.sig)),]

# remove NA


# scale and center normalized counts
mat.sig.scaled <- t(scale(t(mat.sig)))

pdf("scramble_vs_background_sig_heatmap.pdf",
    width = 4,
    height = 5)
Heatmap(mat.sig.scaled, 
        name = "Z-score",
        top_annotation = ha,
        border = TRUE,
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_row_names = FALSE,
        show_column_names = FALSE,
        row_dend_width = unit(20, "mm"))
dev.off()
```

## Arc vs no infect + scramble (PP7)
- Using same genetic background for controls
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")

se.filter$sample.name
se.filter$background

# group experimental and controls
se.filter$group <- c("control",
                     "control",
                     "background",
                     "arc",
                     "arc",
                     "background",
                     "background")

# run deseq
dds <- DESeqDataSet(se.filter, design = ~group)

dds <- DESeq(dds)

# get results
arc_vs_no_infect_scramble <- results(dds, 
                  alpha = 0.05, 
                  contrast = c("group", "arc", "control"))

# get a summary of the amount of significantly DEG 
summary(arc_vs_no_infect_scramble, 
        alpha = 0.05)

# get shrunken fold change
# arc_vs_no_infect_scramble <- lfcShrink(dds = dds,
#                               res = arc_vs_no_infect_scramble,
#                               contrast = c("condition", "actin", "scramble"),
#                               type = "ashr")

# merge
cols <- c("ensembl_gene_id",
          "external_gene_name",
          "entrezgene_id",
          "gene_biotype",
          "chromosome_name",
          "start_position",
          "end_position",
          "percentage_gene_gc_content")

arc_vs_no_infect_scramble <- merge(as.data.frame(rowData(dds)[,cols]),
                         as.data.frame(arc_vs_no_infect_scramble),
                         by = "row.names")


counts <- counts(dds, normalized = TRUE)
colnames(counts) <- dds$sample.description
arc_vs_no_infect_scramble <- merge(arc_vs_no_infect_scramble, 
                         counts,
                         by.x = "Row.names",
                         by = "row.names")

arc_vs_no_infect_scramble$ensembl_gene_id <- NULL
colnames(arc_vs_no_infect_scramble)[1] <- "ensembl_gene_id"

# write results
write.csv(arc_vs_no_infect_scramble, "arc_vs_no_infect_scramble.csv",
          row.names = FALSE)

# p-value distribution histogram
pdf("arc_vs_no_infect_scramble_pvalue_distribution.pdf",
    width = 5,
    height = 5)
hist(arc_vs_no_infect_scramble$pvalue, 
     main = "P-value distribution", 
     xlab = "P-value", 
     ylab = "Frequency")
dev.off()


# ma plot
arc_vs_no_infect_scramble <- arc_vs_no_infect_scramble %>%
  as.data.frame() %>%
  mutate(sig = case_when(log2FoldChange > 0 & arc_vs_no_infect_scramble$padj < 0.05 ~ "sig.up",
                         log2FoldChange < 0 & arc_vs_no_infect_scramble$padj < 0.05 ~ "sig.down",
                         .default = "")
         )

pdf("arc_vs_no_infect_scramble_ma_log2.pdf",
    width = 5,
    height = 5)
ggplot(arc_vs_no_infect_scramble, aes(log2(baseMean + 1), log2FoldChange, color = sig)) +
  geom_point(alpha = 0.5, size = 4) +
  theme_Publication() +
  scale_colour_manual(values = c("grey", "blue", "red")) +
  geom_hline(yintercept= 0, 
             size=1,
             linetype = "dashed") +
  labs(title = "Arc vs Control",
       x = "Average Normalized Log2(counts) + 1",
       y = "Log2(fold change)")
dev.off()

# volcano plot
arc_vs_no_infect_scramble$volcano.label <- ""
arc_vs_no_infect_scramble$volcano.score <- arc_vs_no_infect_scramble$log2FoldChange * -log10(arc_vs_no_infect_scramble$pvalue)
arc_vs_no_infect_scramble <- arc_vs_no_infect_scramble[order(arc_vs_no_infect_scramble$volcano.score, decreasing = TRUE),]
arc_vs_no_infect_scramble$volcano.label[1:10] <- arc_vs_no_infect_scramble$external_gene_name[1:10]

pdf("arc_vs_no_infect_scramble_volcano.pdf",
    width = 5,
    height = 5)
EnhancedVolcano(arc_vs_no_infect_scramble,
                lab = arc_vs_no_infect_scramble$external_gene_name,
                title = expression("RNA-Seq:" ~ italic("Arc") ~ "vs Control"),
                subtitle = NULL,
                caption = "FC > 2; padj < 0.05",
                x = 'log2FoldChange',
                y = 'pvalue',
                pointSize = 2,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = 2,
                pCutoff = 0.05,
                pCutoffCol = 'padj',
                selectLab = arc_vs_no_infect_scramble$volcano.label,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold.italic',
                #labFace = "italic",
                boxedLabels = FALSE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                max.overlaps = Inf,
                gridlines.major = FALSE,
                gridlines.minor = FALSE,
                ylim = c(0, 13),
                xlim = c(-5, 25))
dev.off()

png("arc_vs_no_infect_scramble_volcano.png",
    width = 5,
    height = 5,
    units = "in",
    res = 300)
EnhancedVolcano(arc_vs_no_infect_scramble,
                lab = arc_vs_no_infect_scramble$external_gene_name,
                title = "Arc vs Control",
                subtitle = NULL,
                caption = "FC > 2; padj < 0.05",
                x = 'log2FoldChange',
                y = 'pvalue',
                pointSize = 2,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = 2,
                pCutoff = 0.05,
                pCutoffCol = 'padj',
                selectLab = arc_vs_no_infect_scramble$volcano.label,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = TRUE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                max.overlaps = Inf,
                gridlines.major = FALSE,
                gridlines.minor = FALSE)
dev.off()

# heatmap of significant changes
ha <- HeatmapAnnotation(Condition = subset(colData(dds), group %in% c("control", "arc"))$group)

# get normalized counts
mat <- as.matrix(counts(dds, normalized = TRUE)[,colData(se.filter)$group %in% c("control", "arc")])
colnames(mat) <- paste(dds$condition, dds$replicate)[colData(se.filter)$group %in% c("control", "arc")]
rownames(mat) <- rowData(dds)$external_gene_name

# order by results
mat.order <- mat[match(arc_vs_no_infect_scramble$external_gene_name, rownames(mat)),]

# get significant genes
mat.sig <- mat.order[arc_vs_no_infect_scramble$padj < 0.05 & arc_vs_no_infect_scramble$log2FoldChange > 0,]
mat.sig <- mat.sig[!is.na(row.names(mat.sig)),]

# scale and center normalized counts
mat.sig.scaled <- t(scale(t(mat.sig)))

pdf("arc_vs_no_infect_scramble_sig_heatmap.pdf",
    width = 5,
    height = 8)
Heatmap(mat.sig.scaled, 
        name = "Z-score",
        top_annotation = ha,
        border = TRUE,
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_row_names = TRUE,
        show_column_names = TRUE,
        row_dend_width = unit(20, "mm"),
        row_names_gp = gpar(fontsize = 8))
dev.off()

png("arc_vs_no_infect_scramble_sig_heatmap.png",
    width = 5,
    height = 8,
    units = "in",
    res = 300)
Heatmap(mat.sig.scaled, 
        name = "Z-score",
        top_annotation = ha,
        border = TRUE,
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_row_names = TRUE,
        show_column_names = TRUE,
        row_dend_width = unit(20, "mm"),
        row_names_gp = gpar(fontsize = 8))
dev.off()

make_overlap_heatmap(dds,
                     arc_vs_no_infect_scramble,
                      previous_hits,
                      groups = c("control","arc"),
                      padj_cutoff = 0.05,
                      lfc_direction = "up",
                      lfc_min = 0,
                      sample_label_fun = function(df) paste(df$condition, df$replicate),
                      left_right_ratio = c(2,1),
                      rowname_fontsize = 7,
                      pdf_file = "arc_vs_no_infect_scramble_sig_heatmap.pdf",
                      pdf_width = 5,
                      pdf_height = 8)
```

## Arc vs no infect + scramble (MCP)
- Using same genetic background for controls
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/differential_expression")

se.filter$sample.name
se.filter$background

# group experimental and controls
se.filter$group <- c("control",
                     "background",
                     "control",
                     "arc",
                     "arc",
                     "background",
                     "background")

# run deseq
dds <- DESeqDataSet(se.filter, design = ~group)

dds <- DESeq(dds)

# get results
arc_vs_no_infect_scramble_mcp <- results(dds, 
                  alpha = 0.05, 
                  contrast = c("group", "arc", "control"))

# get a summary of the amount of significantly DEG 
summary(arc_vs_no_infect_scramble_mcp, 
        alpha = 0.05)

# get shrunken fold change
# arc_vs_no_infect_scramble_mcp <- lfcShrink(dds = dds,
#                               res = arc_vs_no_infect_scramble_mcp,
#                               contrast = c("condition", "actin", "scramble"),
#                               type = "ashr")

# merge
cols <- c("ensembl_gene_id",
          "external_gene_name",
          "entrezgene_id",
          "gene_biotype",
          "chromosome_name",
          "start_position",
          "end_position",
          "percentage_gene_gc_content")

arc_vs_no_infect_scramble_mcp <- merge(as.data.frame(rowData(dds)[,cols]),
                         as.data.frame(arc_vs_no_infect_scramble_mcp),
                         by = "row.names")


counts <- counts(dds, normalized = TRUE)
colnames(counts) <- dds$sample.description
arc_vs_no_infect_scramble_mcp <- merge(arc_vs_no_infect_scramble_mcp, 
                         counts,
                         by.x = "Row.names",
                         by = "row.names")

arc_vs_no_infect_scramble_mcp$ensembl_gene_id <- NULL
colnames(arc_vs_no_infect_scramble_mcp)[1] <- "ensembl_gene_id"

# write results
write.csv(arc_vs_no_infect_scramble_mcp, "arc_vs_no_infect_scramble_mcp.csv",
          row.names = FALSE)

# p-value distribution histogram
pdf("arc_vs_no_infect_scramble_mcp_pvalue_distribution.pdf",
    width = 5,
    height = 5)
hist(arc_vs_no_infect_scramble_mcp$pvalue, 
     main = "P-value distribution", 
     xlab = "P-value", 
     ylab = "Frequency")
dev.off()


# ma plot
arc_vs_no_infect_scramble_mcp <- arc_vs_no_infect_scramble_mcp %>%
  as.data.frame() %>%
  mutate(sig = case_when(log2FoldChange > 0 & arc_vs_no_infect_scramble_mcp$padj < 0.05 ~ "sig.up",
                         log2FoldChange < 0 & arc_vs_no_infect_scramble_mcp$padj < 0.05 ~ "sig.down",
                         .default = "")
         )

pdf("arc_vs_no_infect_scramble_mcp_ma_log2.pdf",
    width = 5,
    height = 5)
ggplot(arc_vs_no_infect_scramble_mcp, aes(log2(baseMean + 1), log2FoldChange, color = sig)) +
  geom_point(alpha = 0.5, size = 4) +
  theme_Publication() +
  scale_colour_manual(values = c("grey", "blue", "red")) +
  geom_hline(yintercept= 0, 
             size=1,
             linetype = "dashed") +
  labs(title = "Arc vs Control",
       x = "Average Normalized Log2(counts) + 1",
       y = "Log2(fold change)")
dev.off()

# volcano plot
arc_vs_no_infect_scramble_mcp$volcano.label <- ""
arc_vs_no_infect_scramble_mcp$volcano.score <- arc_vs_no_infect_scramble_mcp$log2FoldChange * -log10(arc_vs_no_infect_scramble_mcp$pvalue)
arc_vs_no_infect_scramble_mcp <- arc_vs_no_infect_scramble_mcp[order(arc_vs_no_infect_scramble_mcp$volcano.score, decreasing = TRUE),]
arc_vs_no_infect_scramble_mcp$volcano.label[1:10] <- arc_vs_no_infect_scramble_mcp$external_gene_name[1:10]

pdf("arc_vs_no_infect_scramble_mcp_volcano.pdf",
    width = 5,
    height = 5)
EnhancedVolcano(arc_vs_no_infect_scramble_mcp,
                lab = arc_vs_no_infect_scramble_mcp$external_gene_name,
                title = "Arc vs Control",
                subtitle = NULL,
                caption = "FC > 2; padj < 0.05",
                x = 'log2FoldChange',
                y = 'pvalue',
                pointSize = 2,
                colAlpha = 0.5,
                legendPosition = "none",
                FCcutoff = 2,
                pCutoff = 0.05,
                pCutoffCol = 'padj',
                selectLab = arc_vs_no_infect_scramble_mcp$volcano.label,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = TRUE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                max.overlaps = Inf,
                gridlines.major = FALSE,
                gridlines.minor = FALSE)
dev.off()



# heatmap of significant changes
ha <- HeatmapAnnotation(Condition = subset(colData(dds), group %in% c("control", "arc"))$group)

# get normalized counts
mat <- as.matrix(counts(dds, normalized = TRUE)[,colData(se.filter)$group %in% c("control", "arc")])
colnames(mat) <- paste(dds$condition, dds$replicate)[colData(se.filter)$group %in% c("control", "arc")]
rownames(mat) <- rowData(dds)$external_gene_name

# order by results
mat.order <- mat[match(arc_vs_no_infect_scramble_mcp$external_gene_name, rownames(mat)),]

# get significant genes
mat.sig <- mat.order[arc_vs_no_infect_scramble_mcp$padj < 0.05 & arc_vs_no_infect_scramble_mcp$log2FoldChange > 0,]
mat.sig <- mat.sig[!is.na(row.names(mat.sig)),]

# scale and center normalized counts
mat.sig.scaled <- t(scale(t(mat.sig)))

pdf("arc_vs_no_infect_scramble_mcp_sig_heatmap.pdf",
    width = 5,
    height = 8)
Heatmap(mat.sig.scaled, 
        name = "Z-score",
        top_annotation = ha,
        border = TRUE,
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_row_names = TRUE,
        show_column_names = TRUE,
        row_dend_width = unit(20, "mm"),
        row_names_gp = gpar(fontsize = 8))
dev.off()
```

# Overlaps

## No infect vs Background & Scramble vs Background
- Together this is the control gene set
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/overlaps")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/overlaps")

x <- list(
"No infect vs Background" = subset(noinfect_vs_background, log2FoldChange > 0 & padj < 0.05)$external_gene_name,
"Scramble vs Background" = subset(scramble_vs_background, log2FoldChange > 0 & padj < 0.05)$external_gene_name)

pdf("noinfect_vs_background_and_scramble_vs_background_overlap.pdf",
    width = 5,
    height = 5)
ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF"),
  stroke_size = 0.5, set_name_size = 4
  )
dev.off()

venn_analysis(x, "noinfect_vs_background_and_scramble_vs_background_overlap.csv")
```

## arc vs no infect + scramble (PP7) & control gene set
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/overlaps")

x <- list(
"arc vs no infect + scramble" = subset(arc_vs_no_infect_scramble, log2FoldChange > 0 & padj < 0.05)$external_gene_name,
"control gene set" = c(subset(scramble_vs_background, log2FoldChange > 0 & padj < 0.05)$external_gene_name,
                       subset(noinfect_vs_background, log2FoldChange > 0 & padj < 0.05)$external_gene_name)
)

pdf("arc_vs_no_infect_scramble_and_control_gene_set_overlap.pdf",
    width = 5,
    height = 5)
ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF"),
  stroke_size = 0.5, set_name_size = 4
  )
dev.off()

venn_analysis(x, "arc_vs_no_infect_scramble_and_control_gene_set_overlap.csv")
```

## arc vs no infect + scramble (PP7) all replicates vs arc vs no infect + scramble (PP7) rep 1 removed
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/overlaps")

arc_vs_no_infect_scramble_all_reps <- read.csv("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc/differential_expression/arc_vs_no_infect_scramble.csv")

x <- list(
"arc vs no infect + scramble - all reps" = subset(arc_vs_no_infect_scramble_all_reps, log2FoldChange > 0 & padj < 0.05)$external_gene_name,
"arc vs no infect + scramble - rep 1 removed" = subset(arc_vs_no_infect_scramble, log2FoldChange > 0 & padj < 0.05)$external_gene_name)

pdf("arc_vs_no_infect_scramble_all_reps_and_arc_vs_no_infect_scramble_rep1_removed_overlap.pdf",
    width = 5,
    height = 5)
ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF"),
  stroke_size = 0.5, set_name_size = 4
  )
dev.off()

venn_analysis(x, "arc_vs_no_infect_scramble_all_reps_and_arc_vs_no_infect_scramble_rep1_removed_overlap.csv")
```

### Correlation
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/correlate")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/correlate")

# scatter plot
df <- left_join(arc_vs_no_infect_scramble_all_reps, 
                arc_vs_no_infect_scramble,
                by = "ensembl_gene_id")
df.plot <- data.frame("All replicates" = df$log2FoldChange.x,
                 "Replicate 1 removed" = df$log2FoldChange.y,
                 genes = df$external_gene_name.x) %>%
  mutate(label_this = All.replicates > 5 & Replicate.1.removed > 5)

pdf("arc_all_reps_vs_rep1_removed.pdf",
    width = 5,
    height = 5)
ggplot(df.plot, aes(x=All.replicates, y=Replicate.1.removed)) +
  geom_hline(yintercept = 0,
             linetype="dashed") +
  geom_vline(xintercept = 0,
             linetype="dashed") +
  geom_point(alpha = 0.25) +
  geom_density_2d() +
  geom_smooth(method=lm) +
  stat_cor() +
  geom_text_repel(
    data = subset(df.plot, label_this),
    aes(label = genes),
    max.overlaps = 20) +
  theme_Publication() +
  labs(title = "Correlation of log2FC",
       x = "All replicates",
       y = "Replicate 1 removed")
dev.off()
```

# Gene set enrichment analysis

## No infect vs Background
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

# get score to use in enrichment
noinfect_vs_background$score <- sign(noinfect_vs_background$log2FoldChange) * -log10(noinfect_vs_background$pvalue)
res.sorted <- noinfect_vs_background[order(noinfect_vs_background$score, decreasing = TRUE),]
res.list <- res.sorted$score
names(res.list) <- res.sorted$entrezgene_id
res.list <- res.list[!is.na(names(res.list))] # remove genes witn no ID
res.list <- res.list[!is.na(res.list)] # remove genes with no score
```

### Biological process
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

bp <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "BP",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
bp.simple <- clusterProfiler::simplify(bp)

write.csv(bp.simple@result, "noinfect_vs_background_go_bp.csv")

pdf("noinfect_vs_background_go_bp.pdf",
    width = 6,
    height = 5)
dotplot(bp.simple, 
        showCategory=10, 
        title = paste("GSE GO BP", "No infect vs Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Moelcular function
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

mf <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "MF",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
mf.simple <- clusterProfiler::simplify(mf)

write.csv(mf.simple@result, "noinfect_vs_background_go_mf.csv")

pdf("noinfect_vs_background_go_mf.pdf",
    width = 6,
    height = 5)
dotplot(mf.simple, 
        showCategory=10, 
        title = paste("GSE GO MF", "No infect vs Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Cellular component
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

cc <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "CC",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
cc.simple <- clusterProfiler::simplify(cc)

write.csv(cc.simple@result, "noinfect_vs_background_go_cc.csv")

pdf("noinfect_vs_background_go_cc.pdf",
    width = 6,
    height = 5)
dotplot(cc.simple, 
        showCategory=10, 
        title = paste("GSE GO CC", "No infect", "vs", "Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

## Scramble vs Background
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

# get score to use in enrichment
scramble_vs_background$score <- sign(scramble_vs_background$log2FoldChange) * -log10(scramble_vs_background$pvalue)
res.sorted <- scramble_vs_background[order(scramble_vs_background$score, decreasing = TRUE),]
res.list <- res.sorted$score
names(res.list) <- res.sorted$entrezgene_id
res.list <- res.list[!is.na(names(res.list))] # remove genes witn no ID
res.list <- res.list[!is.na(res.list)] # remove genes with no score
```

### Biological process
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

bp <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "BP",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
bp.simple <- clusterProfiler::simplify(bp)

write.csv(bp.simple@result, "scramble_vs_background_go_bp.csv")

pdf("scramble_vs_background_go_bp.pdf",
    width = 6,
    height = 5)
dotplot(bp.simple, 
        showCategory=10, 
        title = paste("GSE GO BP", "Scramble vs Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Moelcular function
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

mf <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "MF",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
mf.simple <- clusterProfiler::simplify(mf)

write.csv(mf.simple@result, "scramble_vs_background_go_mf.csv")

pdf("scramble_vs_background_go_mf.pdf",
    width = 6,
    height = 5)
dotplot(mf.simple, 
        showCategory=10, 
        title = paste("GSE GO MF", "Scramble vs Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Cellular component
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

cc <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "CC",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
cc.simple <- clusterProfiler::simplify(cc)

write.csv(cc.simple@result, "scramble_vs_background_go_cc.csv")

pdf("scramble_vs_background_go_cc.pdf",
    width = 6,
    height = 5)
dotplot(cc.simple, 
        showCategory=10, 
        title = paste("GSE GO CC", "Scramble vs Background", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

## Arc vs no infect + scramble (PP7)
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

# get score to use in enrichment
arc_vs_no_infect_scramble$score <- sign(arc_vs_no_infect_scramble$log2FoldChange) * -log10(arc_vs_no_infect_scramble$pvalue)
res.sorted <- arc_vs_no_infect_scramble[order(arc_vs_no_infect_scramble$score, decreasing = TRUE),]
res.list <- res.sorted$score
names(res.list) <- res.sorted$entrezgene_id
res.list <- res.list[!is.na(names(res.list))] # remove genes witn no ID
res.list <- res.list[!is.na(res.list)] # remove genes with no score
```

### Biological process
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

bp <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "BP",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
bp.simple <- clusterProfiler::simplify(bp)

write.csv(bp.simple@result, "arc_vs_no_infect_scramble_go_bp.csv")

pdf("arc_vs_no_infect_scramble_go_bp.pdf",
    width = 6,
    height = 5)
dotplot(bp.simple, 
        showCategory=10, 
        title = paste("GSE GO BP", "Arc vs No infect + Scramble", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Moelcular function
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

mf <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "MF",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
mf.simple <- clusterProfiler::simplify(mf)

write.csv(mf.simple@result, "arc_vs_no_infect_scramble_go_mf.csv")

pdf("arc_vs_no_infect_scramble_go_mf.pdf",
    width = 6,
    height = 5)
dotplot(mf.simple,
        showCategory=10,
        title = paste("GSE GO MF", "Arc vs No infect + Scramble", sep = " "),
        split=".sign",
        x = "NES",
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

### Cellular component
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/functional_enrichment")

cc <- gseGO(geneList     = res.list,
            OrgDb        = org.Mm.eg.db,
            ont          = "CC",
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            verbose      = FALSE)
cc.simple <- clusterProfiler::simplify(cc)

write.csv(cc.simple@result, "arc_vs_no_infect_scramble_go_cc.csv")

pdf("arc_vs_no_infect_scramble_go_cc.pdf",
    width = 6,
    height = 5)
dotplot(cc.simple, 
        showCategory=10, 
        title = paste("GSE GO CC", "Arc vs No infect + Scramble", sep = " "), 
        split=".sign",
        x = "NES", 
        font.size = 10,
        label_format = 50) +
        geom_vline(xintercept = 0, linetype="dotted", color = "#444444", size=1) +
        scale_color_viridis(option = "viridis", direction = 1) +
        labs(fill = "test")
dev.off()
```

# Over representation enrichment analysis

## No infect vs Background

### Upregulated
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")

enrich_up <- enrichGO(gene = subset(noinfect_vs_background, noinfect_vs_background$pvalue < 0.05 & noinfect_vs_background$log2FoldChange > 0)$entrezgene_id,
            universe = as.character(noinfect_vs_background$entrezgene_id),
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, "noinfect_vs_background_enrich_up.csv")

pdf("noinfect_vs_background_enrich_up.pdf",
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "No infect vs Background", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()
```

### Downregulated
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")

enrich_down <- enrichGO(gene = subset(noinfect_vs_background, noinfect_vs_background$pvalue < 0.05 & noinfect_vs_background$log2FoldChange < 0)$entrezgene_id,
            universe = as.character(noinfect_vs_background$entrezgene_id),
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_down.simple <- clusterProfiler::simplify(enrich_down)

write.csv(enrich_down.simple@result, "noinfect_vs_background_enrich_down.csv")

pdf("noinfect_vs_background_enrich_down.pdf",
    width = 10,
    height = 5)
dotplot(enrich_down.simple, 
        showCategory=10, 
        title = paste("GSE GO", "No infect vs Background", "Downregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()
```

## Scramble vs Background

### Upregulated
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")

enrich_up <- enrichGO(gene = subset(scramble_vs_background, scramble_vs_background$pvalue < 0.05 & scramble_vs_background$log2FoldChange > 0)$entrezgene_id,
            universe = as.character(scramble_vs_background$entrezgene_id),
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, "scramble_vs_background_enrich_up.csv")

pdf("scramble_vs_background_enrich_up.pdf",
    width = 10,
    height = 5)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = paste("GSE GO", "Scramble vs Background", "Upregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()
```

### Downregulated
```{r}
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")

enrich_down <- enrichGO(gene = subset(scramble_vs_background, scramble_vs_background$pvalue < 0.05 & scramble_vs_background$log2FoldChange < 0)$entrezgene_id,
            universe = as.character(scramble_vs_background$entrezgene_id),
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_down.simple <- clusterProfiler::simplify(enrich_down)

write.csv(enrich_down.simple@result, "scramble_vs_background_enrich_down.csv")

pdf("scramble_vs_background_enrich_down.pdf",
    width = 10,
    height = 5)
dotplot(enrich_down.simple, 
        showCategory=10, 
        title = paste("GSE GO", "Scramble vs Background", "Downregulated", sep = " "), 
        font.size = 10,
        label_format = 50,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()
```

## Arc vs no infect + scramble (PP7)
```{r}
dir.create("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")
setwd("~/EinsteinMed Dropbox/Ronald Cutler/Vijg-lab/Collaborations/Jackson Rogow/transcriptomics/analysis/with-virus-and-stem-loop/batch-2-arc-rep1-removed/over_representation_enrichment")

enrich_up <- enrichGO(gene = subset(arc_vs_no_infect_scramble, arc_vs_no_infect_scramble$pvalue < 0.05 & arc_vs_no_infect_scramble$log2FoldChange > 0)$entrezgene_id,
            universe = as.character(arc_vs_no_infect_scramble$entrezgene_id),
            OrgDb = org.Mm.eg.db,
            ont = "ALL",
            pool = TRUE,
            keyType = "ENTREZID",
            pvalueCutoff = 0.1,
            readable = TRUE)

enrich_up.simple <- clusterProfiler::simplify(enrich_up)

write.csv(enrich_up.simple@result, "arc_vs_no_infect_scramble_enrich_up.csv")

pdf("arc_vs_no_infect_scramble_enrich_up.pdf",
    width = 10,
    height = 7)
dotplot(enrich_up.simple, 
        showCategory=10, 
        title = "GO: Arc vs Control", 
        font.size = 10,
        label_format = 60,
        split = "ONTOLOGY") +
  facet_wrap(~ONTOLOGY)
dev.off()
```
